Methods paper for erosion

### 2017-08-29 docs/methods/erosion_methods.tex
worked on formulation in methods paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
We consider multiple bodies being eroded by a two-dimensional Stokesian
fluid.  The Stokes paradox is circumvented by using a rounded-off
rectangle that encloses the eroding bodies.  Then, the flow is driven
by a boundary conditions, such as a Hagen-Poiseuille flow, on the
boundary of the outer geometry \todo[inline]{Need a figure here}
\begin{subequations}
\label{eqn:erosionModel}
\begin{gather}
eqn1 \label{eqn:eqn1} \\
eqn2 \label{eqn:eqn2} \\
\pderiv{\xx}{\nn} = -|\tau| \label{eqn:bodyEvolution}
\end{gather}
\end{subequations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Erosion Model} 
\todo[inline]{Governing equations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
The incompressible Stokes equations are
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Gamma.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  The double-layer potential
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal, solves the incompressible Stokes equations for any sufficiently
smooth density function $\eeta$.  However, the double-layer potential can not capture velocity fields
corresponding to rigid body motions~\cite{pow-mir1987}.  A standard
method to complete the double-layer potential is to introduce a
Stokeslet and rotlet due to each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point in the interior of $\Omega_k$, $\rr = \xx -
\cc_k$, and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ can be written in
terms of the completed double-layer potential
\begin{align}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
  \label{eqn:completedDLP}
\end{align}
where the density function, Stokeslets, and rotlets satisfy the
second-kind integral equation
\begin{subequations}
\label{eqn:completed_DLP}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align}
This results in a rank-one null space of
equation~\eqref{eqn:completed_DLP} which we remove by adding the 
rank-one modification $\NN_0[\eeta](\xx) = \int_{\Gamma_0} (\nn(\xx) \times
\nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx \in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
The governing equations~\eqref{eqn:erosionModel} require the shear
stress which is defined on the boundary of each body.  The shear stress
is given by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector.  We note that for the Stokes
equations, because of the no-slip boundary condition on the bodies, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  The velocity field of the
completed double-layer potential involves three terms: the double-layer
potential $\DD$, the Stokeslets $S$, and the rotlets $R$.  To compute
the shear stress, we first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these terms
individually.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
Let us first define the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_k$,
$k=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_k(s,t) \in
\gamma_k$, where $s$ is the arclength and $t$ is time.  Given
$\xx_k(s,t)$, $k=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\dot{\xx}(t) = -|\tau| \nn, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\nn$ is the outward unit normal vector which points away from the
fluid.  The outer wall is formed by rounding off the corners of $[-3,3]
\times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align}
  \uu(\xx) = (1-y^2,0)
\end{align}
on $\Gamma$.  To minimize the boundary effects of the inflow and
outflow, we only place bodies in the center third, $[-1,1] \times
[-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_k$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    Finally, the bodies are eroded according to the shear stress applied
    by the Stokesian fluid.}
\end{figure}

Our erosion model is based on the shear stress, and this model is
discussed in more detail in Section~\ref{sec:erosion}.  We then describe
how the incompressible Stokes equations are solved in $\Omega$ using a
boundary integral equation in Section~\ref{sec:bies}.  Then, in
Section~\ref{sec:qois}, we discuss how the shear stress, pressure, and
drag can all be readily computed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Erosion Model} 
\label{sec:erosion}
\todo[inline]{Describe the erosion model and its justification}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
We use a quasi-static approximation.  That is, at each time step, we
freeze the geometry, solve the incompressible Stokes equations, and then
update the geometry according to the erosion model.  A flow diagram is
illustrated in Figure~\ref{fig:workflow}.

The incompressible Stokes equations are
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Gamma.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  The double-layer potential
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal, solves the incompressible Stokes equations for any sufficiently
smooth density function $\eeta$.  However, the double-layer potential can not capture velocity fields
corresponding to rigid body motions~\cite{pow-mir1987}.  A standard
method to complete the double-layer potential is to introduce a
Stokeslet and rotlet due to each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point in the interior of $\Omega_k$, $\rr = \xx -
\cc_k$, and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ can be written in
terms of the completed double-layer potential
\begin{align}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
  \label{eqn:completedDLP}
\end{align}
where the density function, Stokeslets, and rotlets satisfy the
second-kind integral equation
\begin{subequations}
\label{eqn:completed_DLP}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align}
This results in a rank-one null space of
equation~\eqref{eqn:completed_DLP} which we remove by adding the 
rank-one modification $\NN_0[\eeta](\xx) = \int_{\Gamma_0} (\nn(\xx) \times
\nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx \in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
\label{sec:qois}
The governing equations~\eqref{eqn:erosionModel} require the shear
stress which is defined on the boundary of each body.  The shear stress
is given by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector.  We note that for the Stokes
equations, because of the no-slip boundary condition on the bodies, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  The velocity field of the
completed double-layer potential involves three terms: the double-layer
potential $\DD$, the Stokeslets $S$, and the rotlets $R$.  To compute
the shear stress, we first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these terms
individually.
